Animazing

### 2024-04-14 paper/paper.qmd
Finished draft of model and results

---
title: "My title: still thinking"
subtitle: "Definitely want a subtitle"
author: Allen Uy
thanks: "Code and data are available at: https://github.com/varygx/AnimeAnalysis"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

bind_rows(head(clean_anime, 3), tail(clean_anime, 3)) |>
  select(id, title, rank, start_date, studio, mean, popularity, num_list_users,
         num_scoring_users, adventure, drama) |>
  kable(
    booktabs = TRUE,
    linesep = ""
  ) |> kable_styling(latex_options = "scale_down")
```

@@

### Score Distribution (is there such thing as too many sections?)

@@

### Rank and Popularity

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-date_vs_popularity
#| fig-cap: Relationship Between Date and Popularity

ggplot(clean_anime, aes(x = start_date, y = popularity)) +
  geom_point(color = "blue", size = 0.5, alpha=0.3) +
  labs(x = "Start Date", y = "Popularity") +
  theme_minimal() +
  scale_y_reverse() +
  geom_smooth(method = "lm", color = "red", se = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-date_vs_popularity_modern
#| fig-cap: Relationship Between Date and Popularity in Modern Anime

@@

ggplot(modern_anime, aes(x = start_date, y = popularity)) +
  geom_point(color = "blue", size = 0.5, alpha=0.3) +
  labs(x = "Start Date", y = "Popularity") +
  theme_minimal() +
  scale_y_reverse() +
  geom_smooth(method = "lm", color = "red", se = FALSE)
```

### Date and Popularity

It may seem like the date an anime airs can be a factor for popularity as older anime could have longer exposure over time or newer anime has more coverage but this does not seem to be the case. @fig-date_vs_popularity shows people are less likely to watch very old anime and @fig-date_vs_popularity_modern suggests date is irrelevant to popularity.

@@

mean_fraction <- mean(clean_anime$num_scoring_users / clean_anime$num_list_users)

clean_anime |> ggplot(aes(x=num_scoring_users / num_list_users)) +
  geom_histogram(binwidth = 0.01, color="darkblue", fill="lightblue") +
  labs(x = "Fraction of Scoring Users to List Users", y = "Frequency") +
  annotate("text", x = mean_fraction + 0.1, y = 370, label = paste("Mean Fraction:", round(mean_fraction, 3))) +
  theme_minimal() +
  geom_vline(aes(xintercept=mean_fraction), linetype="dashed", size=1)
```

### Scoring Fraction

On average, 39% of users end up rating an anime on their list according to @fig-scoring_fraction. There is a strong positive correlation between scoring fraction and popularity as seen in @fig-fraction_relationships-1, which is sensible as fans of an anime or any media in general are likely express their opinion on that particular media. Surprisingly in @fig-fraction_relationships-2 there is a positive correlation between scoring fraction and rank. One might expect a normal distribution as people are more likely to voice their opinion on a particularly good or bad piece of work but not so much for average ones. Instead we see a linear relationship scoring fraction and rank.

@@

clean_anime |> ggplot(aes(x = num_scoring_users / num_list_users, y = popularity)) +
  geom_point(color = "blue", size = 0.5, alpha=0.3) +
  labs(x = "Scoring Fraction", y = "Popularity") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  theme_minimal() +
  scale_y_reverse()

clean_anime |> ggplot(aes(x = num_scoring_users / num_list_users, y = rank)) +
  geom_point(color = "blue", size = 0.5, alpha=0.3) +
  labs(x = "Scoring Fraction", y = "Rank") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  theme_minimal() +
  scale_y_reverse()
```

TODO look at studio and genre

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

@@

library(rstanarm)

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

```

# Discussion

@@

# pp_check(first_model) +
#   theme_classic() +
#   theme(legend.position = "bottom")
# 
# posterior_vs_prior(first_model) +
#   theme_minimal() +
#   scale_color_brewer(palette = "Set1") +
#   theme(legend.position = "bottom") +
#   coord_flip()
```

>>>>

---
title: "Animazing"
subtitle: "What was found"
author: Allen Uy
thanks: "Code and data are available at: https://github.com/varygx/AnimeAnalysis"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

bind_rows(head(clean_anime, 3), tail(clean_anime, 3)) |>
  select(id, title, rank, start_date, studio, mean, popularity, num_list_users,
         num_scoring_users, adventure, drama, fraction, days_since_start) |>
  kable(
    booktabs = TRUE,
    linesep = ""
  ) |> kable_styling(latex_options = "scale_down")
```

@@

### Score Distribution {#sec-score-distribution}

@@

### Rank and Popularity {#sec-rank-and-popularity}

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-date_relationships
#| fig-cap: "Correlation to Date"
#| fig-subcap: ["Popularity plot", "Rank plot"]
#| layout-ncol: 2

ggplot(clean_anime, aes(x = start_date, y = popularity)) +
  geom_point(color = "blue", size = 0.5, alpha=0.3) +
  labs(x = "Start Date", y = "Popularity") +
  theme_minimal() +
  scale_y_reverse() +
  geom_smooth(method = "lm", color = "red", se = FALSE)

ggplot(clean_anime, aes(x = start_date, y = rank)) +
  geom_point(color = "blue", size = 0.5, alpha=0.3) +
  labs(x = "Start Date", y = "Rank") +
  theme_minimal() +
  scale_y_reverse() +
  geom_smooth(method = "lm", color = "red", se = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-date_relationships_modern
#| fig-cap: "Correlation to Modern Dates"
#| fig-subcap: ["Popularity plot", "Rank plot"]
#| layout-ncol: 2

@@

ggplot(modern_anime, aes(x = start_date, y = popularity)) +
  geom_point(color = "blue", size = 0.5, alpha=0.3) +
  labs(x = "Start Date", y = "Popularity") +
  theme_minimal() +
  scale_y_reverse() +
  geom_smooth(method = "lm", color = "red", se = FALSE)

ggplot(modern_anime, aes(x = start_date, y = rank)) +
  geom_point(color = "blue", size = 0.5, alpha=0.3) +
  labs(x = "Start Date", y = "Rank") +
  theme_minimal() +
  scale_y_reverse() +
  geom_smooth(method = "lm", color = "red", se = FALSE)
```

### Correlation to Date {#sec-correlation-to-date}

It may seem like the date an anime airs can be a factor for popularity as older anime could have longer exposure over time or newer anime has more coverage but this does not seem to be the case. @fig-date_relationships-1 shows people are less likely to watch very old anime and @fig-date_relationships_modern-1 suggests date is generally irrelevant to popularity but modern anime does have more viewership compared to older anime by the density of the points.

We observe a weaker relationship between date and rating in @fig-date_relationships-2 which suggests viewers are still appreciative of older works. As the amount of anime produced has increased in modern times, @fig-date_relationships_modern-2 suggests that for every 'good' anime there is a 'bad' anime to match it.

@@

mean_fraction <- mean(clean_anime$fraction)

clean_anime |> ggplot(aes(x=fraction)) +
  geom_histogram(binwidth = 0.01, color="darkblue", fill="lightblue") +
  labs(x = "Fraction of Scoring Users to List Users", y = "Frequency") +
  annotate("text", x = mean_fraction + 0.3, y = 370, label = paste("Mean Fraction:", round(mean_fraction, 3))) +
  theme_minimal() +
  geom_vline(aes(xintercept=mean_fraction), linetype="dashed", size=1)
```

### Scoring Fraction {#sec-scoring-fraction}

On average, 39% of users end up rating an anime on their list according to @fig-scoring_fraction. There is a strong positive correlation between scoring fraction and popularity as seen in @fig-fraction_relationships-1, which is sensible as fans of an anime or any media in general are likely express their opinion on that particular media. Surprisingly in @fig-fraction_relationships-2 there is a positive correlation between scoring fraction and rank. One might expect a normal distribution as people are more likely to voice their opinion on a particularly good or bad piece of work but not so much for average ones. Instead we see a linear relationship between scoring fraction and rank.

@@

clean_anime |> ggplot(aes(x = fraction, y = popularity)) +
  geom_point(color = "blue", size = 0.5, alpha=0.3) +
  labs(x = "Scoring Fraction", y = "Popularity") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  theme_minimal() +
  scale_y_reverse()

clean_anime |> ggplot(aes(x = fraction, y = rank)) +
  geom_point(color = "blue", size = 0.5, alpha=0.3) +
  labs(x = "Scoring Fraction", y = "Rank") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  theme_minimal() +
  scale_y_reverse()
```

# Model {#sec-model}

Here we briefly describe the Bayesian analysis model used to investigate the user score of anime on MAL. Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model Set-up

From the preliminary analysis conducted in @sec-data, we build a multiple linear regression model to predict the rating of an anime based on quantitative statistics. Define $y_i$ as the user score of the anime $i$. Then $\text{fraction}_i$ is the fraction of users who have rated the anime $i$ to members of the anime $i$, $\text{popularity}_i$ is the total members of anime $i$, and $\text{startDate}_i$ is the number of days from the start date of anime $i$ to the current date.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \text{fraction} + \beta_2 \times \text{popularity} + \beta_3 \times \text{startDate}\\
\beta_0 &\sim \mbox{Normal}(0, 1.6) \\
\beta_1 &\sim \mbox{Normal}(0, 13) \\
\beta_2 &\sim \mbox{Normal}(0, 3.6 \times 10^{-4}) \\
\beta_3 &\sim \mbox{Normal}(0, 6.3 \times 10^{-6}) \\
\sigma &\sim \mbox{Exponential}(1.6)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @citeRstanarm. We use the default priors from `rstanarm` which are Normal distributions and apply autoscaling for better results.

### Model Justification

We expect a positive relationship between the score of the anime and scoring fraction as seen in @sec-scoring-fraction. This is likely to be the most important factor in our model. We expect a very minor to no relationship between the score of the anime and the start date of an anime as seen in @sec-correlation-to-date. As such this coefficient is expected to be close to zero. Finally, we expect a minor positive relationship between the score of an anime and its popularity as seen in @sec-rank-and-popularity. This coefficient will likely be close to zero and smaller than start date, as an anime can have millions of members, but should be a better indicator than start date.

# Results {#sec-results}

Our results are summarized in @tbl-modelresults. Our model observes a score intercept of 6.02, which means when an anime has been out for 0 days, has 0 members and thus 0 scoring fraction, it will likely have a user score of 6.34. This is slightly nonsensical as an anime had to have a positive scoring fraction and the minimum amount of scoring users to be ranked within the top 10000. When considering only the mean scoring fraction in @sec-scoring-fraction of 0.39, our model predicts a user score of $6.34 + 1.55\times0.39=6.94$, which is close to the mean user score of 6.91 seen in @sec-score-distribution. The start date coefficient is extremely small indicating the date of an anime has no impact on its score. For a 10 year old anime the model expects the score to decrease by $-1.605 \times 10^{-5} \times (10 \times 365)=-0.06$, which can be considered a neglible amount. Finally, for every 1 million members of an anime, its score increases by $7.77 \times 10^{-7}\times10^6=0.78$ .

@@

library("rstanarm")
model <- readRDS(here::here("models/model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of score based on scoring fraction, popularity, and start date"
#| warning: false

library("modelsummary")

my_fmt <- list(
  intercept = list(fmt = "%.2f"),
  fraction = list(fmt = "%.2f"),
  days_since_start = list(fmt = "%#.2e"),
  num_list_users = list(fmt = "%#.2e")
)

table <- modelsummary(
  model,
  statistic = "mad",
  fmt = my_fmt,
  output = "kableExtra"
)

table |> kable_styling(latex_options = "scale_down")
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelpredictions
#| tbl-cap: "Comparing predictions of model to actual scores"
#| warning: false

predictions <- clean_anime |>
  filter(!is.na(fraction) & !is.na(num_list_users) & !is.na(days_since_start)) |>
  mutate(predicted = predict(model, .)) |>
  select(title, rank, popularity, mean, num_list_users, fraction, start_date, days_since_start, predicted)

bind_rows(head(predictions, 5), tail(predictions, 5)) |>
  select(id, title, rank, start_date, studio, mean, popularity, num_list_users,
         num_scoring_users, adventure, drama, fraction, days_since_start) |>
  kable(
    booktabs = TRUE,
    linesep = ""
  ) |> kable_styling(latex_options = "scale_down")
```

@tbl-modelpredictions shows the predictions of our model compared to the actual scores of the top 5 and bottom 5 anime in the dataset. Some observations that should be noted are:

1.  The score intercept is higher than the score of lower ranked anime
2.  The model is harsh on highly ranked anime that are not yet popular

# Discussion {#sec-discussion}

@@

# analysis_data <- clean_anime

pp_check(model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```
